From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-07-29 WOLFHPC15/article/concept.tex
fix compiler 29_07

Let introduce some concepts used in this paper. First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain. A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure). For example, a structured mesh is illustrated in Figure~\ref{fig:ex}, each cell contains four nodes and four edges. Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes, for example. In Figure~\ref{fig:ex} two mesh domains are used, the center of the cells (in red) and the edge on the horizontal axe (in blue on the left and right of each cell). A \emph{data} is a quantity to simulate or a temporary data to perform computations. A data is mapped onto a given mesh domain. For example, in Figure~\ref{fig:ex}, data $A$ and $B$ are mapped onto the center of cells, while $C$ is mapped onto the edges.

\begin{figure}[!h]\begin{center}
  \resizebox{7cm}{!}{\includegraphics{./images/stencil.pdf}}
  \caption{Examples of a stencil computations}
  \label{fig:ex}
\end{center}\end{figure}

The first stencil computation in Figure~\ref{fig:ex} computes $B$ using $A$, while the second one computes $B$ using $C$. Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result. The \emph{computation domain} is a subset of elements to compute on the output data. In other words a computation domain is a subset of a mesh domain. For example, in the first stencil, the computation of $B$ is performed on plain elements, while dotted elements are not computed. On the other hand, on the second computation, the computation domain of $B$ is exactly equal to the cell mesh domain. The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.

Finally, a stencil computation is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data. The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}. For example, the stencil shape of the first computation in Figure~\ref{fig:ex} contains direct neighbors on the right, left, top and bottom. Sometimes, the neighborhood shape can also access a different mesh domain, as for example in the second computation example. Actually, in this computation, the neighborhood contains edges on the left and on the right of the cell. As a result, we could imagine for example that the numerical expression of the first example is
\begin{equation*} 
B(x,y) = A(x+1,y)+A(x-1,y)+A(x,y+1)+A(x,y-1),
\end{equation*}
and that the numerical expression of the second example is
\begin{equation*} 
B(x,y) = B(x,y)+C(x1,y1)+C(x1+1,y1).
\end{equation*}

>>>>

Let introduce some concepts used in this paper. First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain. A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure). For example, a structured mesh is illustrated in Figure~\ref{fig:mesh}, each cell contains four nodes and four edges. Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes, for example. In Figure~\ref{fig:mesh} two mesh domains are illustrated, the center of the cells (in red) and edges on the horizontal axe (in blue on left and right of each cell). A \emph{data} is a quantity to simulate or a temporary data to perform computations. A data is mapped onto a given mesh domain. For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while in Figure~\ref{fig:ex2} $C$ is mapped onto the edges.

\begin{figure}
\begin{center}
\subfloat[Mesh and mesh domains\label{fig:mesh}]{
\resizebox{7cm}{!}{\includegraphics{./images/mesh.pdf}}
}
\hspace{10pt}
\subfloat[First stencil example\label{fig:ex1}]{
\resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
}
\hspace{10pt}
\subfloat[Second stencil example\label{fig:ex2}]{
\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
}
\end{center}
\caption{Stencil examples}
\label{fig:gspmsp}
\end{figure}

% \begin{figure}[!h]\begin{center}
%   \resizebox{7cm}{!}{\includegraphics{./images/stencil.pdf}}
%   \caption{Examples of a stencil computations}
%   \label{fig:ex}
% \end{center}\end{figure}

The first stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the second one in Figure~\ref{fig:ex2} computes $A$ using $C$. Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result. The \emph{computation domain} is a subset of elements to compute on the output data. In other words a computation domain is a subset of a mesh domain. For example, in the first stencil, the computation of $A$ is performed on plain elements, while dotted elements are not computed. On the other hand, on the second computation, the computation domain of $A$ is exactly equal to the cell mesh domain. The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.

Finally, a stencil computation is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data. The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}. For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom. Sometimes, the neighborhood shape can also access a different mesh domain, as for example in Figure~\ref{fig:ex2}. Actually, in this computation, the neighborhood contains edges on the left and on the right of the cell. As an example, the numerical expression of the first example could be
\begin{equation*} 
B(x,y) = A(x+1,y)+A(x-1,y)+A(x,y+1)+A(x,y-1),
\end{equation*}
and the numerical expression of the second example could be
\begin{equation*} 
B(x,y) = B(x,y)+C(x1,y1)+C(x1+1,y1).
\end{equation*}
